### 2018-02-27 2_impl.tex

\chapter{Implementation}
\label{section:impl}

This chapter describes the architecture of the tool \texttt{mousquitaires} ...

language: java

\section{Program Requirements}
\label{section:impl:requirements}

- stability (tests)
- scalability (new features of language, new models, new tasks for a program -- e.g.!)
- transparency
- efficiency

\section{Program Components}
\label{section:impl:comp}

Big view

\subsection{C11 to YTree parser}
\label{section:impl:comp:ytree}

- The language-dependent syntax tree:
        - for now it's the C subset language which I called 'Cmin'; as a base, I used the C11 grammar from ANTLR github repository, then I simplified it a lot, cutting off many unnecessary C syntax features and making it more convenient for parsing. When developing the Cmin language, I kept in mind C elements that are necessary for processing the linux kernel code, though for now not the whole grammar element described in file 'Cmin.g4' are being implemented;
        - later I am going to add the litmus grammar as well;
        - in future, it will be not a problem to add any new C-like language;

- The language-independent abstract syntax tree (aliased 'Ytree', where 'Y' resembles branching of the tree):
        - all tree nodes in my code are prefixed with 'Y', see tentative (yet almost complete) class hierarchy in picture 'YEntity.png';
        - this AST contains very basic language elements according to the C execution model (statements and expressions);
        - converting the language-dependent syntax tree to the language-independent syntax tree is performed by Visitor pattern (e.g., for Cmin->Ytree conversion is made by 'CminToYtreeConverterVisitor')
        - minor changes are performed by converting to ytree representation: desugaring the target code, etc.

\subsection{YTree to XGraph event converter}
\label{section:impl:comp:xgraph}

- Then, the AST is being interpreted and converted to event-based representation (aliased 'Xrepr' for eXecution representation):
        - more low-level code representation (or high-level assembly);
        - I try to keep this representation close to the one you described in your papers: basic load \& store events, branching events, fence events;
        - this representation is being implementing these days, I've just started doing it (see current class hierarchy in the picture 'XEntity.png');

- After we acquired the event-based representation, we can perform some modifications/simplifications/optimisations on it (separately, allowing user to manage them):
        - converting to SSA form as one of necessary steps before encoding;
        - (more? -- I'm not thinking about it yet);

\subsection{XGraph to ZFormula (SMT) encoder}
\label{section:impl:comp:zformula}

- Then, this modified event-representation is being encoded to SMT formula and sent to the solver.

\section{Optimisations}

... in each stage