### 2018-04-07 4_input.tex
Complete the data-flow encoding.

\chapter{The input language}
\label{ch:impl}

the old input language syntax

new one iwth changes highlighted