### 2018-05-14 99_summary.tex
finish impl tentatively. add draft for eval.

\chapter{Summary}
\label{ch:summary}

enumerate goals (from Task specification) in past tense + justification

contribution

  architecture:  framework

  new encoding: more laconic (compare!) + actually show size

  kernel???