### 2018-05-18 99_summary.tex
cleanup a bit

\chapter{Summary}
\label{ch:summary}

enumerate goals (from Task specification) in past tense + justification

contribution

  architecture:  framework

  new encoding: more laconic (compare!) + actually show size

  kernel???
  %http://diy.inria.fr/doc/litmus.html#klitmus

\section{Comparison with \porthos{}}

\subsection{New features}

enumerate extensions of the grammar

... (?)