### 2018-05-21 99_summary.tex

\chapter{Summary}
\label{ch:summary}

enumerate goals (from Task specification) in past tense + justification

contributions:

  architecture:  framework; restate requirements

  new encoding: more laconic? + actually show num of clauses

  %kernel http://diy.inria.fr/doc/litmus.html#klitmus

  enumerate extensions of the grammar

  ... (?)

\begin{itemize}
\item Improved performance
\item kernel partial support
\end{itemize}

Future work