### 2018-05-24 99_summary.tex

\chapter{Conclusion}
\label{ch:summary}

\section{Solved tasks and contributions}

During the work on this thesis, we solved the following engineering and scientific tasks:

\begin{enumerate}[label=\Roman*.]
\item
Firstly, we studied existing weak memory model-aware analysis frameworks and tools to gather deep understanding of the problem (Sections~\ref{ch:intro:related} and~\ref{ch:enc:mc} and Chapter~\ref{ch:wmm}).

\item
Then, we explored the existing implementation of the proof-of-concept tool \porthos[1] in order to understand components that need to be redesigned in order to facilitate the extension of the input language.

  \begin{enumerate}[label=\roman*.,leftmargin=\parindent]
  \item The first stumbling block in the input language extension was the \porthos[1] input language parser, which performs the full semantic analysis of the program.
  Although this works for a small toy language, the rich and expressive language such as C requires several stages of analysis before the compilation stage.
  For handling that, we implemented the processing units that operate on the \textit{pre-compilation} stage discussed in Section~\ref{ch:impl:proc:x-pre-compiler}.

  \item Secondly, as \porthos[2] determines the \textit{kind (shared or local) of each variable} on the pre-compilation phase, the shared variables can be accessed in arbitrary code location (comparing to \porthos[1] which determines the type of a variable syntactically, depending on the operator of function that uses the variable).

  \item In addition, \porthos[2] supports some \textit{new syntactic constructions} of the C language: \texttt{break} and \texttt{continue} jumps, function invocations, multiple-variable declarations (as `\texttt{int a, b=2, z;}'), arbitrary expressions (corresponding to the C standard), and others (see Section~\ref{ch:impl:input}).

  \item As the C language supports unconditional \texttt{goto}-jumps that can produce an arbitrary control-flow graph, which cannot be supported solely with the program AST, therefore we needed to compile the AST to the low-level hardware-agnostic program representation \xgraph{}.
  For that, we implemented the \textit{full C interpreter} discussed in Section~\ref{ch:impl:proc:x-compiler}.

  \item Originally, the control-flow instructions were encoded directly into the SMT-formula~\cite{Porthos17a}.
  In contrast, \porthos[2] encodes the low-level \xgraph{} representation, which can have arbitrary graph structure.
  For encoding the \xgraph{} into the SMT-formula, we apply the \textit{new control-flow encoding scheme} that in general follows the one proposed in~\cite[Chapter 5.1.2]{heljanko2008unfoldings} (see Section~\ref{ch:enc:bmc:cf}).
  As the new encoding scheme does not add new variables to every control-flow instruction, the number of variables in the result SMT-formula is expected to be smaller.

  \item For ease the process of adding support for new domain-specific functions (for example, the Kernel-specific atomic write function \texttt{atomic\_store}), we implemented the \textit{invocation hooking mechanism}, which allows to modify the compilation process without changing the interpreter.
  The invocation hooks serve as a knowledge base for the program domain, that are to be extended in future.
  We illustrate this mechanism with the basic support for Linux kernel litmus tests.

  \item Since the new tool compiles the program AST to the \xgraph{} before encoding it into the SMT-formula, we decided to change the \textit{unrolling algorithm} from the simple unrolling all loops $k$ times (where $k$ is the user-specified unrolling bound) to the DFS-based algorithm that explores all possible states that the program can reach within $k$ steps (see discussion on the unrolling in Sections~\ref{ch:impl:proc:x-unroll} and~\ref{ch:eval:show:compil}).
  Although the new algorithm produces considerably more events than the old algorithm and thus the requires more time for the program domain encoding, we claim it to be complete.

  \end{enumerate}

\item 
As the tests show, the overhead of the new architecture is negligible, therefore we consider the applied architectural changes as acceptable.

\end{enumerate}

Thus, the new tool \porthos[2] with easily extensible knowledge base that allow to support new domain-specific functions can be considered as a generalised framework for SMT-based memory model-aware analysis.

%enumerate goals (from Task specification) in past tense + justification

\section{Limitations and future work}

- improve the encoding: an abstraction over Z

- typing: const-size arrays + structures and so on
  -only basic datatypes

- hooks : flexible. but still in Java, not as separate module (as it requires turing-complete language to be expressive enough) => extension of the knowledge base for supporting new platforms / languages

- new modes: as it was discussed <before>, works only in the litmus test-mode. May work with complete projects. But here: 
  - memory issues (we store same graph + back edges + memoised queries ==> multiple instances.);
    - + will need the memory-guarding module
  - ability to encode partially (once the

-only one mode litmus-test intra proc

- testing with new solvers